---
title: "CellDMC - new version"
author: "Sarah Lee"
date: "2022-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(EpiDISH)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(ggsci)
```

```{r}
# tumor purity
purity <- read.csv("~/TumorPurity.csv")
```

# Subset
```{r}
meta_sub <- meta_cytmod[meta_cytmod$OX_BS == 0,]
```

# Reformat
```{r}
# Tumor vs normal
meta_sub$tissue <- ifelse(meta_sub$Tumor_Categories %in% "Non-Tumor", "Non-Tumor", "Tumor")

table(meta_sub$tissue)
```

```{r}
# Transpose
prop <- read.csv('~/singlecell_proportions.csv')
prop_t <- as.data.frame(t(prop))
```


```{r}
# Reformat metadata
# Add onto meta data
meta_full2 <- merge(meta_sub, prop_t, 
                   by.x = "SampleName_Old", by.y = "row.names")


```{r}
# Add tumor purity
meta_full3 <- merge(meta_full2, purity, by.x = "SampleName", by.y = "X")
```

 Make sure everything is in the same order
```{r}
meta_reord <- meta_full3[match(colnames(final_5hmC), meta_full3$SampleName), ]
all(meta_reord$SampleName == colnames(final_5hmC))
```

```{r}
colnames(final_5hmC) <- meta_reord$SampleName_Old
colnames(final_5mC) <- meta_reord$SampleName_Old
```


```{r}
prop_reord <- prop_t[match(colnames(final_5hmC), rownames(prop_t)), ]
all(rownames(prop_reord) == colnames(final_5hmC))
```

# Make into matrix
```{r}
mC_mtx <- as.matrix(final_5mC)
hmC_mtx <- as.matrix(final_5hmC)

# Add some cell types together
prop_reord$PROG <- rowSums(prop_reord[,c("OPC", "RGC", "NSC", "UBC")])
prop_mtx <- as.matrix(prop_reord)
```

```{r}
# Subset proportions
prop_sub <- prop_mtx[, colnames(prop_mtx) %in% c("PROG", "NEU")]
```

```{r}
which(colSums(model.matrix(~ prop_sub + meta_reord$tissue:prop_sub)[, -1]) == 0)
```

```{r}
rowSums(prop_sub)
```

```{r}
# Reformat grade
meta_reord$grade <- as.factor(ifelse(meta_reord$WHO_GRADE %in% NA, 0, meta_reord$WHO_GRADE))
```

# Run CellDMC
#### BY TUMOR TYPE
```{r}
atc_meta <- meta_reord[meta_reord$tumortype_refact %in% c("Astrocytoma", "Non-Tumor"), ]
atc_mC_mtx <- mC_mtx[, atc_meta$SampleName_Old]
atc_hmC_mtx <- hmC_mtx[, atc_meta$SampleName_Old]
atc_prop <- prop_sub[atc_meta$SampleName_Old, ]

all(colnames(atc_mC_mtx) == atc_meta$SampleName_Old)
all(colnames(atc_mC_mtx) == rownames(atc_prop))
```

```{r}
emb_meta <- meta_reord[meta_reord$tumortype_refact %in% c("Embryonal", "Non-Tumor"), ]
emb_mC_mtx <- mC_mtx[, emb_meta$SampleName_Old]
emb_hmC_mtx <- hmC_mtx[, emb_meta$SampleName_Old]
emb_prop <- prop_sub[emb_meta$SampleName_Old, ]

all(colnames(emb_mC_mtx) == emb_meta$SampleName_Old)
```

```{r}
epn_meta <- meta_reord[meta_reord$tumortype_refact %in% c("Ependymoma", "Non-Tumor"), ]
epn_mC_mtx <- mC_mtx[, epn_meta$SampleName_Old]
epn_hmC_mtx <- hmC_mtx[, epn_meta$SampleName_Old]
epn_prop <- prop_sub[epn_meta$SampleName_Old, ]

all(colnames(epn_mC_mtx) == epn_meta$SampleName_Old)
```

```{r}
gnn_meta <- meta_reord[meta_reord$tumortype_refact %in% c("GNN", "Non-Tumor"), ]
gnn_mC_mtx <- mC_mtx[, gnn_meta$SampleName_Old]
gnn_hmC_mtx <- hmC_mtx[, gnn_meta$SampleName_Old]
gnn_prop <- prop_sub[gnn_meta$SampleName_Old, ]

all(colnames(gnn_mC_mtx) == gnn_meta$SampleName_Old)
```

```{r}
atc_dmc_res <- CellDMC(beta.m = atc_mC_mtx, 
                   pheno.v = atc_meta$tissue, 
                   frac.m = atc_prop,
                   #cov.mod = model.matrix(~ Sex + Age_at_Diagnosis, data = atc_meta),
                   mc.cores = 4)

atc_dhmc_res <- CellDMC(beta.m = atc_hmC_mtx, 
                   pheno.v = atc_meta$tissue, 
                   frac.m = atc_prop,
                   #cov.mod = model.matrix(~ Sex + Age_at_Diagnosis, data = atc_meta),
                   mc.cores = 4)

```

```{r}
emb_dmc_res <- CellDMC(beta.m = emb_mC_mtx, 
                   pheno.v = emb_meta$tissue, 
                   frac.m = emb_prop,
                   #cov.mod = model.matrix(~ Sex + Age_at_Diagnosis, data = emb_meta),
                   mc.cores = 4)

emb_dhmc_res <- CellDMC(beta.m = emb_hmC_mtx, 
                   pheno.v = emb_meta$tissue, 
                   frac.m = emb_prop,
                   #cov.mod = model.matrix(~ Sex + Age_at_Diagnosis, data = emb_meta),
                   mc.cores = 4)
```

```{r}
epn_dmc_res <- CellDMC(beta.m = epn_mC_mtx, 
                   pheno.v = epn_meta$tissue, 
                   frac.m = epn_prop,
                   #cov.mod = model.matrix(~ Sex + Age_at_Diagnosis, data = epn_meta),
                   mc.cores = 4)

epn_dhmc_res <- CellDMC(beta.m = epn_hmC_mtx, 
                   pheno.v = epn_meta$tissue, 
                   frac.m = epn_prop,
                   #cov.mod = model.matrix(~ Sex + Age_at_Diagnosis, data = epn_meta),
                   mc.cores = 4)

```

```{r}
gnn_dmc_res <- CellDMC(beta.m = gnn_mC_mtx, 
                   pheno.v = gnn_meta$tissue, 
                   frac.m = gnn_prop,
                   #cov.mod = model.matrix(~ Sex + Age_at_Diagnosis, data = gnn_meta),
                   mc.cores = 4)

gnn_dhmc_res <- CellDMC(beta.m = gnn_hmC_mtx, 
                   pheno.v = gnn_meta$tissue, 
                   frac.m = gnn_prop,
                   #cov.mod = model.matrix(~ Sex + Age_at_Diagnosis, data = gnn_meta),
                   mc.cores = 4)
```





